TMR: Detecting Structure in Graphical Data

### 2021-04-08 final-report/final-report.tex
Implement uniform prior on b

\documentclass[11pt]{article}

@@

%packages
\usepackage[margin=0.6in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage{parskip}
\usepackage{fancyvrb}
\usepackage{bbm}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{multicol}
\usepackage{titling}

@@

\begin{definition}
	\label{defn:sbm}
	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities. We define the vector $p = [p_1, p_2 \dots p_k]^T$ to be the prior on the k-communities. Each vertex $v \in \Vcal = \{1, 2 \dots n\}$ has a community label $X_v \in \{1, 2 \dots n\}$. Let $W$ be a symmetric $k \times k$ matrix with entries in $[0,1]$ called the connectivity matrix. The pair $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if X is an $n \times 1$ vector with each component independently drawn from the community prior $X_v \sim p$ and $\Gcal$ is an $n$-vertex graph where each vertex-pair $(i, j)$ is connected with probability $\prob(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other edges. Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which contains all vertices belonging to community $i$.
\end{definition}

@@

For our inference goals we wish to sample block memberships $b$ according to the posterior $p(b | \Gcal)$. Peixoto \cite{Peixoto-MCMC} proposes a Monte-Carlo method for generating these samples. We simply

>>>>

\documentclass[]{article}

@@

%packages
\usepackage[margin=0.6in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage{parskip}
\usepackage{fancyvrb}
\usepackage{bbm}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{multicol}
\usepackage{titling}
\usepackage{tikz}

@@

\begin{definition}[Canonical SBM]
	\label{defn:sbm}
	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities. We define the vector $p = [p_1, p_2 \dots p_k]^T$ to be the prior on the k-communities. Each vertex $v \in \Vcal = \{1, 2 \dots n\}$ has a community label $X_v \in \{1, 2 \dots n\}$. Let $W$ be a symmetric $k \times k$ matrix with entries in $[0,1]$ called the connectivity matrix. The pair $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if X is an $n \times 1$ vector with each component independently drawn from the community prior $X_v \sim p$ and $\Gcal$ is an $n$-vertex graph where each vertex-pair $(i, j)$ is connected with probability $\prob(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other edges. Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which contains all vertices belonging to community $i$.
\end{definition}

@@

For our inference goals we wish to sample block memberships $b$ according to the posterior $p(b | \Gcal)$. Peixoto \cite{Peixoto-MCMC} proposes a Monte-Carlo method for generating these samples. However, he proposes a slight modification of the Stochastic Block Model. The so called microcanonical formulation. In contrast to the canonical version where edge parameters are obeyed only on expectation, the microcanonical model has these constraints obeyed with certainty. The definition follows from \cite{Peixoto-Bayesian-Microcanonical}.

\begin{definition}[Microcanonical SBM]
	\label{defn:microcan-sbm}
	Let $N \in \Integers^{+}$ denote the number of vertices in our graph. The block memberships are denoted by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of nonempty blocks. Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$). We restrict our analysis to undirected graphs so $e$ is symmetric. For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
	%
	\begin{equation}
		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
	\end{equation}
	%
	Where edges are placed at random but respecting the constraints imposed by $e$ and $b$. The additional parameters $N$ and $B$ are omitted as they are inferred from the shape ofs of $b$ and $e$. If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i=1}^{N} \sum_{j=1}^{N} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$. However, this formulation does not tolerate high degree variation within blocks as is typical of real-world data. We therefore introduce the degree-corrected SBM (DC-SBM) which has an additional parameter $k$ which is a vector of length $N$ encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
	%
	\begin{equation}
		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
	\end{equation}
	%
	This imposes the additional constraint that $k_i = \sum_{j=1}^{N} A_{ij}$. In what follows, we will always assume the degree-corrected model unless otherwise specified.

\end{definition}

\begin{figure}[!h]
	\begin{tikzpicture}
		\draw (0, 0) circle (2);
		\draw (2, 0) -> (4, 0);
		\draw (6, 0) circle (2);
		\draw (8, 0) -> (10, 0);
		\draw (12, 0) circle (2);
	\end{tikzpicture}
\end{figure}
